---
title: "p8105_hw3_kk3154"
author: "Kristen King"
date: "10/20/2021"
output: github_document
---

## Problem 1 - Exploring Instacart Data

Loading data from p8105.datasets:

```{r}
library(p8105.datasets)
data("instacart")
```

1.1 Describe the dataset, including size and structure of the data, key variables, and giving illustrative examples of observations.

```{r}

```

1.2 How many aisles are there, and which aisles are the most items ordered from?

```{r}

```

1.3 Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}

```

1.4 Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}

```

1.5 Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}

```

## Problem 2 - Cleaning and Exploring BRFSS Data

Loading the BRFSS data:

```{r}
data("brfss_smart2010")
```

2.1 Data cleaning:

* format the data to use appropriate variable name
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}

```

2.2 In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}

```

2.3 Construct a dataset that is limited to `Excellent` responses, and contains year, state, and a variable that averages the `data_value` across locations within a state. 

```{r}

```

2.4 Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the `geom_line` geometry and `group` aesthetic will help).

```{r}

```

2.5 Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}

```

## Problem 3 - Accelerometer data from a patient with congestive heart failure

35 days of accelerometer data where activity.* variables are activity counts for each minute of a 24-hour day starting at midnight.

3.1 Loading, tidying, and wrangling data.

The final dataset includes all originally observed variables and values, has useful variable names, includes a weekday vs. weekend variable, and encodes data with reasonable variable classes.

```{r}

```

3.2 Describing the resulting dataset (e.g. what variables exist, how many observations, etc.):



3.3 Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}

```

3.4 Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}

```










